Sparse spectral and $p$-finite element methods for partial differential equations on disk-slices and trapeziums

### 2019-05-17 papers/halfdisk/halfdisk.tex
Include new image. Update text.

% FIGURES
\begin{figure}
	\begin{subfigure}[t]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-poisson}
	\caption{}
	\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}[t]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-helmholtz}
	\caption{}
	\label{fig:solution-helmholtz}
	\end{subfigure}
	\begin{subfigure}[t]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-biharmonic}
	\caption{}
	\label{fig:solution-biharmonic}
	\end{subfigure}
	\centering
    	\caption{(a): The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. (b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 200$. (c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
        \label{fig:solution}
\end{figure}

@@

\bstodoinline{N=100, k=200 in the example isn't converging to a solution, so need a new fig (do N=200, k=20). Add in solution block norms fig  (for N=200, k = 20).}

@@

We can tackle the problem as follows. Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopii$ OP basis up to degree $N$ by $\mathbf{f}$. Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule above. We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}. We can denote the resulting operator acting on coefficients in the $\bighdopoo$ space by $V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top)$. In matrix-vector notation, our system hence becomes:
\begin{align*}
    (\laplacewii + k^2 T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
\end{align*}
which can be solved to find $\mathbf{u}$. We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example. In Figure \ref{fig:solution-helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with  $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$. In Figure \ref{fig:solutionblocknorms-helmholtz} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogenous variable-coefficient Helmholtz equation with $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns. The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution. \bstodo{Correct the explanation of what we observe} We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence. For the final Gaussian bump example, we see that we achieve spectral convergence.

>>>>

% FIGURES
\begin{figure}
	\begin{subfigure}[t]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-poisson}
	\caption{}
	\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}[t]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-helmholtz-k=100}
	\caption{}
	\label{fig:solution-helmholtz}
	\end{subfigure}
	\begin{subfigure}[t]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-biharmonic}
	\caption{}
	\label{fig:solution-biharmonic}
	\end{subfigure}
	\centering
    	\caption{(a): The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. (b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = 1 - (3(x-1)^2 + 5y^2)$ and $k = 100$. (c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
        \label{fig:solution}
\end{figure}

@@

@@

We can tackle the problem as follows. Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bighdopii$ OP basis up to degree $N$ by $\mathbf{f}$. Since $f$ is known, we can obtain  the coefficients $\mathbf{f}$ using the quadrature rule above. We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matrices a la the procedure introduced in \cite{olver2019triangle}. We can denote the resulting operator acting on coefficients in the $\bighdopoo$ space by $V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top)$. In matrix-vector notation, our system hence becomes:
\begin{align*}
    (\laplacewii + k^2 T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
\end{align*}
which can be solved to find $\mathbf{u}$. We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example. In Figure \ref{fig:solution-helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with $k=100$, $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ and $f(x,y) = xy e^x$. In Figure \ref{fig:solutionblocknorms-helmoltz} we show the norms of each block of calculated coefficients of the approximation for four right-hand sides of the inhomogenous variable-coefficient Helmholtz equation with $k=20$ and $v(x,y) = (1 - (3(x-1)^2 + 5y^2))$ using N = 200, that is, 20,301 unknowns. The rate of decay in the coefficients is a proxy for the rate of convergence of the computed solution. \bstodo{Correct the explanation of what we observe} We see that we achieve algebraic convergence for the first three examples, noting that for right hand sides that vanish at the corners of our half disk ($x=0, \: y = \pm 1$) we see faster convergence. For the final Gaussian bump example, we see that we achieve spectral convergence.
